---
title: "ESM 204: Homework 4 (A Climate Change Model)"
author: "David Segan, Grace Brofman, & Jordan Isken"
date: "5/21/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)
```

```{r}
# Install libraries as necessary
library(tidyverse)
library(janitor)
library(scales) # for formatting as a percent
library(purrr)
library(pracma) # fzero function is like excel solver
```


```{r}
# -------------------
# Create all of our functions
# -------------------

# Create a function for temperature increase relative to year 0
# t = year
# T = temp increase after year 100
temp <- function(t, T) 
{
  out = pmin(t*T/100, T)
  # pmin works but min does not when looking at a full df. this is because pmin looks at a vector while min looks at all values so would also report back 0
  return(out)
}

# Function for fraction of economic activity retained in a year
econ <- function(beta, temp)
{
  out = exp(-1*beta*(temp^2))
  return(out)
}

# Function for consumption
C <- function(econ, g, t)
{
  out = econ*exp(g*t)
  return(out)
}

# Function for society's utility from consumption
U <- function(C, n)
{
  out = (C^(1-n))/(1-n)
  return(out)
}

```


```{r}
# -----------------
# Set up base case parameters
# -----------------

l = 0.005
n = 0.5
g = 0.01
beta = 0.05

# discount rate
r = l + n*g
```


### 1. Plots

```{r}
# First we need to make data frames that include t from 0 to 200 and also the BAU_T values for each of our difference scenarios

# Assign values
t <- 0:200
T_0 <- 0
T_2 <- 2
T_8 <- 8
T_4.4 <- 4.4

# Create data frames
T_0_df <- data.frame(year = t,
                 temp_increase = T_0,
                 temp = temp(t, T_0)) %>% 
  mutate(fraction_economy_retained0 = econ(beta,temp)) %>% 
  mutate(consumption0 = C(fraction_economy_retained0, g, year)) %>% 
  mutate(utility0 = U(consumption0, n)) %>% 
  mutate(pv_utility0 = utility0/((1+r)^t))

T_2_df <- data.frame(year = t,
                 temp_increase = T_2,
                 temp = temp(t, T_2)) %>% 
  mutate(fraction_economy_retained2 = econ(beta,temp)) %>% 
  mutate(consumption2 = C(fraction_economy_retained2, g, year)) %>% 
  mutate(utility2 = U(consumption2, n))

T_8_df <- data.frame(year = t,
                 temp_increase = T_8,
                 temp = temp(t, T_8)) %>% 
  mutate(fraction_economy_retained8 = econ(beta,temp)) %>% 
  mutate(consumption8 = C(fraction_economy_retained8, g, year)) %>% 
  mutate(utility8 = U(consumption8, n))

```

##### a. Temperature over time

```{r}
ggplot() +
  geom_line(data = T_0_df,
            aes(x = year,
                y = temp,
                color = "No climate change"),
            size = 1) +
  geom_line(data = T_2_df,
            aes(x = year,
                y = temp,
                color = "Modest climate change"),
            size = 1)+
  geom_line(data = T_8_df,
            aes(x = year,
                y = temp,
                color = "Extreme climate change"),
            size = 1) +
  scale_color_manual(values = c("red", "orange", "blue")) +
  theme_minimal() +
  theme(legend.position = c(0.2,.85)) +
  labs(title =,
       x = "Years from year 0",
       y = "Temperature Increase",
       color = "Scenario") 
```


##### b. Consumption over time

```{r}

ggplot() +
  geom_line(data = T_0_df,
            aes(x = year,
                y = consumption0,
                color = "No climate change"),
            size = 1)+
  geom_line(data = T_2_df,
            aes(x = year,
                y = consumption2,
                color = "Modest climate change"),
            size = 1)+
  geom_line(data = T_8_df,
            aes(x = year,
                y = consumption8,
                color = "Extreme climate change"),
            size = 1) +
  scale_color_manual(values = c("red", "orange", "blue")) +
  theme_minimal() +
  theme(legend.position = c(0.2,.85)) +
  labs(x = "Year from year 0",
       y = "Consumption",
       color = "Scenario")

```

##### c. Undiscounted utility over time

```{r}

ggplot() +
  geom_line(data = T_0_df,
            aes(x = year,
                y = utility0,
                color = "No climate change"),
            size = 1)+
  geom_line(data = T_2_df,
            aes(x = year,
                y = utility2,
                color = "Modest climate change"),
            size = 1)+
  geom_line(data = T_8_df,
            aes(x = year,
                y = utility8,
                color = "Extreme climate change"),
            size = 1) +
  scale_color_manual(values = c("red", "orange", "blue")) +
  theme_minimal() +
  theme(legend.position = c(0.2,.85)) +
  labs(x = "Year from year 0",
       y = "Utility",
       color = "Scenario")


```

### 2. Analysis

##### a. Climate change leads to 4.4 degree increase in temperature

```{r}

# Present value / discounted utility 
# Since we want to find the total present value. We have to sum up the present values of each year's utility. To do this let's create another data frame

T_4.4_df <- data.frame(year = t,
                 temp_increase = T_4.4,
                 temp = temp(t, T_4.4)) %>% 
  mutate(fraction_economy_retained4.4 = econ(beta,temp)) %>% 
  mutate(consumption4.4 = C(fraction_economy_retained4.4, g, year)) %>% 
  mutate(utility4.4 = U(consumption4.4, n)) %>% 
  # add column for discounted utility
  mutate(pv_utility4.4 = utility4.4/((1+r)^t))

# Let's graph this to check

# ggplot()+ 
#   geom_line(data = T_4.4_df, 
#        aes(x = year, y = pv_utility))
# 
# ggplot()+ 
#   geom_line(data = T_0_df, 
#        aes(x = year, y = utility))
# 
# ggplot()+ 
#   geom_line(data = T_0_df, 
#        aes(x = year, y = pv_utility))

# Yeah, seems a lot less than undiscounted utility

# Now, how do we sum up all those values....
pv_U_with_cc <- sum(T_4.4_df$pv_utility4.4)
pv_U_without_cc <- sum(T_0_df$pv_utility0)

# Percent difference
pct_loss_L <- (pv_U_with_cc-pv_U_without_cc)/pv_U_without_cc

```

- Present value utility with climate change: **`r round(pv_U_with_cc, 2)`**

- Present value utility without climate change: **`r round(pv_U_without_cc, 2)`**

- Percentage change in present value utility from climate change (L): **`r percent(pct_loss_L)`**


##### b. Sensitivity analysis

```{r}
# Increase parameters by 10%
T_new <- 1.1*T_4.4
g_new <- 1.1*g
n_new <- 1.1*n
beta_new <- 1.1*beta

T_new_df <- data.frame(year = t,
                 temp_increase = T_new,
                 temp = temp(t, T_new)) %>% 
  mutate(fraction_economy_retained = econ(beta,temp)) %>% 
  mutate(consumption = C(fraction_economy_retained, g, year)) %>% 
  mutate(utility = U(consumption, n)) %>% 
  mutate(pv_utility = utility/((1+r)^t))

# Sum PV
pv_U_T_new <- sum(T_new_df$pv_utility)

# New percent difference (L)
pct_loss_L_T_new <- (pv_U_T_new-pv_U_without_cc)/pv_U_without_cc

# Percent change in L

pct_change_L_T_new <- (pct_loss_L_T_new -pct_loss_L)/pct_loss_L

```

###### *Increase T by 10% to `r T_new`:*
- New L: **`r percent(pct_loss_L_T_new)`**
- % change in L from no climate change scenario: **`r percent(pct_change_L_T_new)`**



```{r}
# Increase g by 10%:

# new discount rate
r_g_new = l + n*g_new

g_new_df <- data.frame(year = t,
                 temp_increase = T_4.4,
                 temp = temp(t, T_4.4)) %>% 
  mutate(fraction_economy_retained = econ(beta,temp)) %>% 
  mutate(consumption = C(fraction_economy_retained, g_new, year)) %>%
  mutate(utility = U(consumption, n)) %>% 
  mutate(pv_utility = utility/((1+r)^t))

# Sum PV
pv_U_g_new <- sum(g_new_df$pv_utility)

# New percent difference (L)
pct_loss_L_g_new <- (pv_U_g_new-pv_U_without_cc)/pv_U_without_cc

# Percent change in L

pct_change_L_g_new <- (pct_loss_L_g_new -pct_loss_L)/pct_loss_L

```

###### *Increase g (the consumption growth rate) by 10% to `r g_new`:*
- New L: **`r percent(pct_loss_L_g_new)`**
- % change in L from no climate change scenario: **`r percent(pct_change_L_g_new)`**



```{r}
# Increase n by 10%

# new discount rate
r_n_new = l + n_new*g

n_new_df <- data.frame(year = t,
                 temp_increase = T_4.4,
                 temp = temp(t, T_4.4)) %>% 
  mutate(fraction_economy_retained = econ(beta,temp)) %>% 
  mutate(consumption = C(fraction_economy_retained, g, year)) %>% 
  mutate(utility = U(consumption, n_new)) %>% 
  mutate(pv_utility = utility/((1+r_n_new)^t))

# Sum PV
pv_U_n_new <- sum(n_new_df$pv_utility)

# New percent difference (L)
pct_loss_L_n_new <- (pv_U_n_new-pv_U_without_cc)/pv_U_without_cc

# Percent change in L
pct_change_L_n_new <- (pct_loss_L_n_new -pct_loss_L)/pct_loss_L

```

###### *Increase n (the elasticity of marginal utility of consumption with respect to income) by 10% to `r n_new`:*
- New L: **`r percent(pct_loss_L_n_new)`**
- % change in L from no climate change scenario: **`r percent(pct_change_L_n_new)`**


```{r}
# Increase beta by 10%

beta_new_df <- data.frame(year = t,
                 temp_increase = T_4.4,
                 temp = temp(t, T_4.4)) %>% 
  mutate(fraction_economy_retained = econ(beta_new,temp)) %>% 
  mutate(consumption = C(fraction_economy_retained, g, year)) %>% 
  mutate(utility = U(consumption, n)) %>% 
  mutate(pv_utility = utility/((1+r)^t))

# Sum PV
pv_U_beta_new <- sum(beta_new_df$pv_utility)

# New percent difference (L)
pct_loss_L_beta_new <- (pv_U_beta_new-pv_U_without_cc)/pv_U_without_cc

# Percent change in L
pct_change_L_beta_new <- (pct_loss_L_beta_new -pct_loss_L)/pct_loss_L

```

###### *Increase beta by 10% to `r beta_new`:*
- New L: **`r percent(pct_loss_L_beta_new)`**
- % change in L from no climate change scenario: **`r percent(pct_change_L_beta_new)`**

##### c. Maximum fraction (theta star) of consumption that society would be willing to give up each year to prevent climate change

```{r}
# we need a new consumption amount that leads to a present value utility without climate change equals the present value of utility with 4.4 degrees of climate change

# let's join the data frame for no climate change and the one with 4.4 climate change

theta_df <- full_join(T_4.4_df, T_0_df, by = "year") %>% 
  mutate(C_difference = consumption0 - consumption4.4)


# theta_func <- function(theta)
# {
#   # 
#   out = (((theta)^(1-n))*pv_U_without_cc) - pv_U_with_cc
#   return(out)
# }
# 
# theta_star <- fzero(fun = theta_func, x = 0)$x
# 
# percent(theta_star)
```


##### d. What is theta* (see above) if T is uncertain? 

```{r}
# Uncertainty parameters for T and their associated probabilities

T_u_low <- 2
T_u_low_p <- 0.2
T_u_med <- 4
T_u_med_p <- 0.5
T_u_high <- 6
T_u_high_p <- 0.3
```

