---
title: "ESM 204: Homework 4 (A Climate Change Model)"
author: "David Segan, Grace Brofman, & Jordan Isken"
date: "5/21/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)
```

```{r}
# Install libraries as necessary
library(tidyverse)
library(janitor)
library(scales) # for formatting as a percent
library(purrr) # map functions are like for loops
library(pracma) # fzero function is like excel solver
```


```{r}
# -------------------
# Create all of our functions
# -------------------

# Create a function for temperature increase relative to year 0
# t = year
# T = temp increase after year 100
temp <- function(t, T) 
{
  out = pmin(t*T/100, T)
  # pmin works but min does not when looking at a full df. this is because pmin looks at a vector while min looks at all values so would also report back 0
  return(out)
}

# Function for fraction of economic activity retained in a year
econ <- function(beta, temp)
{
  out = exp(-1*beta*(temp^2))
  return(out)
}

# Function for consumption
C <- function(econ, g, t)
{
  out = econ*exp(g*t)
  return(out)
}

# Function for society's utility from consumption
U <- function(C, n)
{
  out = (C^(1-n))/(1-n)
  return(out)
}

```


```{r}
# -----------------
# Set up base case parameters
# -----------------

l = 0.005
n = 0.5
g = 0.01
beta = 0.05

# discount rate
r = l + n*g
```


### 1. Plots

```{r}
# First we need to make data frames that include t from 0 to 200 and also the BAU_T values for each of our difference scenarios

# Assign values
t <- 0:200
T_0 <- 0
T_2 <- 2
T_8 <- 8
T_4.4 <- 4.4

# Create data frames
T_0_df <- data.frame(year = t,
                 temp_increase = T_0,
                 temp = temp(t, T_0)) %>% 
  mutate(fraction_economy_retained0 = econ(beta,temp)) %>% 
  mutate(consumption0 = C(fraction_economy_retained0, g, year)) %>% 
  mutate(utility0 = U(consumption0, n)) %>% 
  mutate(pv_utility0 = utility0/((1+r)^t))

T_2_df <- data.frame(year = t,
                 temp_increase = T_2,
                 temp = temp(t, T_2)) %>% 
  mutate(fraction_economy_retained2 = econ(beta,temp)) %>% 
  mutate(consumption2 = C(fraction_economy_retained2, g, year)) %>% 
  mutate(utility2 = U(consumption2, n))

T_8_df <- data.frame(year = t,
                 temp_increase = T_8,
                 temp = temp(t, T_8)) %>% 
  mutate(fraction_economy_retained8 = econ(beta,temp)) %>% 
  mutate(consumption8 = C(fraction_economy_retained8, g, year)) %>% 
  mutate(utility8 = U(consumption8, n))

```

##### a. Temperature over time

```{r}
ggplot() +
  geom_line(data = T_0_df,
            aes(x = year,
                y = temp,
                color = "No climate change"),
            size = 1) +
  geom_line(data = T_2_df,
            aes(x = year,
                y = temp,
                color = "Modest climate change"),
            size = 1)+
  geom_line(data = T_8_df,
            aes(x = year,
                y = temp,
                color = "Extreme climate change"),
            size = 1) +
  scale_color_manual(values = c("red", "orange", "blue")) +
  theme_minimal() +
  theme(legend.position = c(0.2,.85)) +
  labs(title =,
       x = "Years from year 0",
       y = "Temperature Increase",
       color = "Scenario") 
```


##### b. Consumption over time

```{r}

ggplot() +
  geom_line(data = T_0_df,
            aes(x = year,
                y = consumption0,
                color = "No climate change"),
            size = 1)+
  geom_line(data = T_2_df,
            aes(x = year,
                y = consumption2,
                color = "Modest climate change"),
            size = 1)+
  geom_line(data = T_8_df,
            aes(x = year,
                y = consumption8,
                color = "Extreme climate change"),
            size = 1) +
  scale_color_manual(values = c("red", "orange", "blue")) +
  theme_minimal() +
  theme(legend.position = c(0.2,.85)) +
  labs(x = "Year from year 0",
       y = "Consumption",
       color = "Scenario")

```

##### c. Undiscounted utility over time

```{r}

ggplot() +
  geom_line(data = T_0_df,
            aes(x = year,
                y = utility0,
                color = "No climate change"),
            size = 1)+
  geom_line(data = T_2_df,
            aes(x = year,
                y = utility2,
                color = "Modest climate change"),
            size = 1)+
  geom_line(data = T_8_df,
            aes(x = year,
                y = utility8,
                color = "Extreme climate change"),
            size = 1) +
  scale_color_manual(values = c("red", "orange", "blue")) +
  theme_minimal() +
  theme(legend.position = c(0.2,.85)) +
  labs(x = "Year from year 0",
       y = "Utility",
       color = "Scenario")


```

### 2. Analysis

##### a. Climate change leads to 4.4 degree increase in temperature

```{r}

# Present value / discounted utility 
# Since we want to find the total present value. We have to sum up the present values of each year's utility. To do this let's create another data frame

T_4.4_df <- data.frame(year = t,
                 temp_increase = T_4.4,
                 temp = temp(t, T_4.4)) %>% 
  mutate(fraction_economy_retained4.4 = econ(beta,temp)) %>% 
  mutate(consumption4.4 = C(fraction_economy_retained4.4, g, year)) %>% 
  mutate(utility4.4 = U(consumption4.4, n)) %>% 
  # add column for discounted utility
  mutate(pv_utility4.4 = utility4.4/((1+r)^t))

# Let's graph this to check

# ggplot()+ 
#   geom_line(data = T_4.4_df, 
#        aes(x = year, y = pv_utility))
# 
# ggplot()+ 
#   geom_line(data = T_0_df, 
#        aes(x = year, y = utility))
# 
# ggplot()+ 
#   geom_line(data = T_0_df, 
#        aes(x = year, y = pv_utility))

# Yeah, seems a lot less than undiscounted utility

# Now, how do we sum up all those values....
pv_U_with_cc <- sum(T_4.4_df$pv_utility4.4)
pv_U_without_cc <- sum(T_0_df$pv_utility0)

# Percent difference
pct_loss_L <- (pv_U_with_cc-pv_U_without_cc)/pv_U_without_cc

```

- Present value utility with climate change: **`r round(pv_U_with_cc, 2)`**

- Present value utility without climate change: **`r round(pv_U_without_cc, 2)`**

- Percentage change in present value utility from climate change (L): **`r percent(pct_loss_L)`**


##### b. Sensitivity analysis

```{r}
# Increase parameters by 10%
T_new <- 1.1*T_4.4
g_new <- 1.1*g
n_new <- 1.1*n
beta_new <- 1.1*beta

T_new_df <- data.frame(year = t,
                 temp_increase = T_new,
                 temp = temp(t, T_new)) %>% 
  mutate(fraction_economy_retained = econ(beta,temp)) %>% 
  mutate(consumption = C(fraction_economy_retained, g, year)) %>% 
  mutate(utility = U(consumption, n)) %>% 
  mutate(pv_utility = utility/((1+r)^t))

# Sum PV
pv_U_T_new <- sum(T_new_df$pv_utility)

# New percent difference (L)
pct_loss_L_T_new <- (pv_U_T_new-pv_U_without_cc)/pv_U_without_cc

# Percent change in L

pct_change_L_T_new <- (pct_loss_L_T_new -pct_loss_L)/pct_loss_L

```

###### *Increase T by 10% to `r T_new`:*
- New L: **`r percent(pct_loss_L_T_new)`**
- % change in L from no climate change scenario: **`r percent(pct_change_L_T_new)`**



```{r}
# Increase g by 10%:

# new discount rate
r_g_new = l + n*g_new

g_new_df <- data.frame(year = t,
                 temp_increase = T_4.4,
                 temp = temp(t, T_4.4)) %>% 
  mutate(fraction_economy_retained = econ(beta,temp)) %>% 
  mutate(consumption = C(fraction_economy_retained, g_new, year)) %>%
  mutate(utility = U(consumption, n)) %>% 
  mutate(pv_utility = utility/((1+r)^t))

# Sum PV
pv_U_g_new <- sum(g_new_df$pv_utility)

# New percent difference (L)
pct_loss_L_g_new <- (pv_U_g_new-pv_U_without_cc)/pv_U_without_cc

# Percent change in L

pct_change_L_g_new <- (pct_loss_L_g_new -pct_loss_L)/pct_loss_L

```

###### *Increase g (the consumption growth rate) by 10% to `r g_new`:*
- New L: **`r percent(pct_loss_L_g_new)`**
- % change in L from no climate change scenario: **`r percent(pct_change_L_g_new)`**



```{r}
# Increase n by 10%

# new discount rate
r_n_new = l + n_new*g

n_new_df <- data.frame(year = t,
                 temp_increase = T_4.4,
                 temp = temp(t, T_4.4)) %>% 
  mutate(fraction_economy_retained = econ(beta,temp)) %>% 
  mutate(consumption = C(fraction_economy_retained, g, year)) %>% 
  mutate(utility = U(consumption, n_new)) %>% 
  mutate(pv_utility = utility/((1+r_n_new)^t))

# Sum PV
pv_U_n_new <- sum(n_new_df$pv_utility)

# New percent difference (L)
pct_loss_L_n_new <- (pv_U_n_new-pv_U_without_cc)/pv_U_without_cc

# Percent change in L
pct_change_L_n_new <- (pct_loss_L_n_new -pct_loss_L)/pct_loss_L

```

###### *Increase n (the elasticity of marginal utility of consumption with respect to income) by 10% to `r n_new`:*
- New L: **`r percent(pct_loss_L_n_new)`**
- % change in L from no climate change scenario: **`r percent(pct_change_L_n_new)`**


```{r}
# Increase beta by 10%

beta_new_df <- data.frame(year = t,
                 temp_increase = T_4.4,
                 temp = temp(t, T_4.4)) %>% 
  mutate(fraction_economy_retained = econ(beta_new,temp)) %>% 
  mutate(consumption = C(fraction_economy_retained, g, year)) %>% 
  mutate(utility = U(consumption, n)) %>% 
  mutate(pv_utility = utility/((1+r)^t))

# Sum PV
pv_U_beta_new <- sum(beta_new_df$pv_utility)

# New percent difference (L)
pct_loss_L_beta_new <- (pv_U_beta_new-pv_U_without_cc)/pv_U_without_cc

# Percent change in L
pct_change_L_beta_new <- (pct_loss_L_beta_new -pct_loss_L)/pct_loss_L

```

###### *Increase beta by 10% to `r beta_new`:*
- New L: **`r percent(pct_loss_L_beta_new)`**
- % change in L from no climate change scenario: **`r percent(pct_change_L_beta_new)`**

##### c. Maximum fraction (theta star) of consumption that society would be willing to give up each year to prevent climate change assuming T = 4.4

```{r}
# we need a new consumption amount that leads to a present value utility without climate change equals the present value of utility with 4.4 degrees of climate change

# let's join the data frame for no climate change and the one with 4.4 climate change

# theta_df <- full_join(T_4.4_df, T_0_df, by = "year") %>%
#   mutate(C_difference = consumption0 - consumption4.4)

# theta_func <- function(theta)
# {
#   #
#   out = (((theta)^(1-n))*pv_U_without_cc) - pv_U_with_cc
#   return(out)
# }
# 
# theta_star <- fzero(fun = theta_func, x = 0)$x
# 
# percent(theta_star)

#----------------------------------
#theta function: in this function we apply 1-theta first to a vector of the consumption values we calculated in part one for the no climate change scenario, then we apply that to our utility function, then to the PV function, and take a sum. This summed total is then subtracted from the PV value we calculated in part 2a.
# ----------------------------------

theta_function <- function(theta)pv_U_with_cc-sum((((1-theta)*T_0_df$consumption0)^(1-n)/(1-n))/(1+r)^t)

#use uniroot to find theta where difference in PV = 0 between climate change scenarios
#"tol =" refers to the number of decimal places
theta_2c_output <- uniroot(theta_function, interval = c(0,1), tol = .0001)

```

Theta star: **`r round(theta_2c_output$root,4)`**

##### d. What is theta star if T is uncertain? 

```{r}
# Uncertainty parameters for T and their associated probabilities

T_u_low <- 2
T_u_low_p <- 0.2
T_u_med <- 4
T_u_med_p <- 0.5
T_u_high <- 6
T_u_high_p <- 0.3

# probability distribution
uncertain_T <- 2*.2 + 4*.5 + 6*.3

output_2d <- function(t,T)
{
time_span <- 100
l = 0.005
n = 0.5
g = 0.01
beta = 0.05
# discount rate
r = l + n*g

temp = pmin(t*T/time_span, T)
econ = exp(-1*beta*(temp^2))
C = econ*exp(g*t)
U = (C^(1-n))/(1-n)
PV = (U/(1+r)^t)

return(data.frame(time=t,
                  T_change=T,
                    temperature = temp,
                    econ = econ,
                    consumption = C,
                  utility = U,
                  present_value = PV))
}

uncertain_T_df <- purrr::map2_dfr(rep(seq(from=0,to=200,by=1), 3), c(rep(2, 201), rep(4, 201), rep(6, 201)), output_2d)

# new data frames for summed PVs using theta function from 2c
uncertain_2_df <- uncertain_T_df %>%
  filter(T_change == 2) %>%
  select(consumption)
uncertain_4_df <- uncertain_T_df %>%
  filter(T_change == 4) %>%
  select(consumption)
uncertain_6_df <- uncertain_T_df %>%
  filter(T_change == 6) %>%
  select(consumption)

# find summed PVs of the three different scenarios, multiplied by probabilities, and then added all together to get the PV of utility under uncertainty.
uncertain_pv_g <-
    ((sum((((uncertain_2_df$consumption))^(1-n)/(1-n))/(1+r)^t))*.2 +
    (sum((((uncertain_4_df$consumption))^(1-n)/(1-n))/(1+r)^t))*.5 +
    (sum((((uncertain_6_df$consumption))^(1-n)/(1-n))/(1+r)^t))*.3)


#theta function: same as in 2c except now we're comparing with the PV for the uncertainty scenario instead of the 4.4 scenario.
theta_function_2d <- function(theta){uncertain_pv_g-sum(((((1-theta)*T_0_df$consumption0))^(1-n)/(1-n))/(1+r)^t)}

#find theta where difference in PV = 0 between climate change scenarios
theta_2d <- uniroot(theta_function_2d, interval = c(0,1), tol = .0001)

```

Theta star under uncertainty: **`r round(theta_2d$root, 4)`**